Weighted Model Counting with Conditional Measures

### 2020-07-21 paper.tex
Edit

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

\begin{itemize}
\item Thesis: many important computational problems are solved by encoding them
  as WMC. But if the weight function is extended to `conditional probabilities',
  the problem becomes easier because current approaches need workarounds in
  order to be independent.
\item Shorter thesis: many important problems are encoded as WMC, but they can
  be encoded in a better way if we allow for conditional probabilities.
\item The claim behind this paper is that allowing for conditional probabilities
  in the context of weighted model counting seems to be a good idea. Bayesian
  networks and ADDMC are only particular examples. This should also work with
  Cachet.
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Contributions
  \begin{itemize}
  \item WMC defines a measure over a BA.
  \item WMC with weights on literals imposes an independence assumption.
    (Measures are `slightly' more expressive than WMC with weights on models
    because they apply to non-atomic BAs.)
  \item A BA can be augmented with new literals in order to support any measure.
  \item (Maybe) a lower bound on the number of new literals needed in order to
    support any measure.
  \item This results in a smaller problem for WMC algorithms (w.r.t. both the
    number of literals and the length of the theory) and thus faster inference.
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item[F2] The paper at this stage is very technical---the danger is that WMC/SRL
  people may not be able to follow it and so would be hard to get accepted.
  Without clear target audience, [???] get work accepted. My main high level
  suggestion is that let us tease apart what you have and see if a story
  emerges. That is, let us attempt to write a paper {\bf with examples} and
  see whether with significant motivation, we have a story emerging. Below,
  sample text needed to adequately motivate [???] for WMC/SRL community.
\end{itemize}

@@

\begin{itemize}
\item[F2] Explain with examples: models $=$ elements [atoms] of algebra.
\item Perhaps reorder the section of preliminaries into paragraphs, i.e., a
  paragraph for order, for homomorphisms, etc. This would take up less space.
\item Make up my mind about $a, b$ vs. $x, y$ and stick to it (maybe $x, y$?).
\item Terminology: `with generating set $S$' $\to$ `over $S$'.
\item Notation: if $L$ denotes literals, then it doesn't denote a generating
  set. Literals should be $U$.
\end{itemize}

@@

\begin{definition}[\cite{givant2008introduction}]
  Let $\mathbf{A}$ and $\mathbf{B}$ be BAs. A \emph{(Boolean) homomorphism} from
  $\mathbf{A}$ to $\mathbf{B}$ is a map $f\colon \mathbf{A} \to \mathbf{B}$ such
  that:
  \begin{itemize}
  \item $f(x \land y) = f(x) \land f(y)$,
  \item $f(x \lor y) = f(x) \lor f(y)$,
  \item $f(\neg x) = \neg f(x)$
  \end{itemize}
  for all $x, y \in \mathbf{A}$.
\end{definition}

\begin{lemma}[Homomorphisms preserve order
  \cite{givant2008introduction}] \label{lemma:homomorphisms_and_order}
  Let $f\colon \mathbf{A} \to \mathbf{B}$ be a homomorphism between two BAs
  $\mathbf{A}$ and $\mathbf{B}$. Then, for any $x, y \in \mathbf{A}$, if $x \le
  y$, then $f(x) \le f(y)$.
\end{lemma}

@@

\begin{definition}[\cite{koppelberg1989handbook}]
  Let $S$ be a set, and let $\mathbf{B}$ be a BA. Then $\mathbf{B}$ is a
  \emph{free BA over $S$} if there is a map $S \to \mathbf{B}$ such that for any
  BA $\mathbf{C}$ and map $S \to \mathbf{C}$, there is a unique homomorphism
  $\mathbf{B} \to \mathbf{C}$ that makes
  \[
    \begin{tikzcd}
      S \ar[rd] \ar[r] & \mathbf{B} \ar[d,dashed] \\
      & \mathbf{C}.
    \end{tikzcd}
  \]
  commute. A BA $\mathbf{B}$ is \emph{free} if $S$ exists.
\end{definition}

\begin{lemma}[\cite{sikorski1969boolean}] \label{lemma:finite_and_free}
  A finite BA is free if and only if it has $2^{2^n}$ elements for some $n \in
  \mathbb{N}$. It then has $2^n$ atoms and $n$ generators.
\end{lemma}

@@

\begin{itemize}
\item Any function $2^U \to \mathbb{R}_{\ge 0}$ induces a measure on
  $2^{2^U}$---the full Boolean algebra. This measure is not necessarily
  probabilistic.
\item[F2] Need to explain how WMC and NWMC connects to standard definitions of
  WMC and $\Pr(\phi \mid e) = \WMC(\phi \land e)/\WMC(e)$.
\item[F2] You need to explain what precisely these mean in logic and models
  and weight functions are usually defined and understood.
\item Be careful about mentioning ideals, filters, and quotients.
\item Note that we're going to use logical notation for BAs. Give an example of
  the two notations.
\end{itemize}

\begin{definition} \label{def:algebra_from_logic}
  Let $\mathcal{L}$ be a propositional (or first-order) logic, and let
  $\Delta$ be a theory in $\mathcal{L}$. We can define an equivalence
  relation on formulas in $\mathcal{L}$ as
  \[
    \alpha \sim \beta \quad \text{if and only if} \quad \Delta \vdash \alpha
    \leftrightarrow \beta
  \]
  for all $\alpha, \beta \in \mathcal{L}$. Let $[\alpha]$ denote the equivalence
  class of $\alpha \in \mathcal{L}$ with respect to $\sim$. We can then let
  $B(\Delta) = \{ [\alpha] \mid \alpha \in \mathcal{L} \}$ and define the
  structure of a BA on $B(\Delta)$ as
  \begin{align*}
    [\alpha] \lor [\beta] &= [\alpha \lor \beta], \\
    [\alpha] \land [\beta] &= [\alpha \land \beta], \\
    \neg[\alpha] &= [\neg\alpha], \\
    1 &= [\alpha \to \alpha], \\
    0 &= [\alpha \land \neg\alpha]
  \end{align*}
  for all $\alpha, \beta \in \mathcal{L}$. Then $B(\Delta)$ is the
  \emph{Lindenbaum-Tarski algebra} of $\Delta$
  \cite{koppelberg1989handbook,tarski1983logic}.
\end{definition}

\begin{example} \label{example:construction}
  Let $\mathcal{L}$ be a propositional logic with $p$ and $q$ as its only atoms.
  Then $L = \{ p, q, \neg p, \neg q \}$ is its set of literals. Let $w : L \to
  \mathbb{R}_{\ge 0}$ be the \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \\
    w(\neg p) = 0.7, \\
    w(q) = 0.2, \\
    w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of $\Delta$ is
  then
  \[
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \]

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$ and then take a quotient with respect to
  either the filter generated by $p$ or the ideal\footnote{More details on these
    concepts can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate}
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
\todo[inline,caption={}]{
  \begin{itemize}
  \item Why is Step 1 always possible?
  \item Clarify what $B(L)$ means and whether $B(\Delta)$ is even necessary.
  \item Find a reference for the set/subset thing.
  \item This should be replaced with inner sums (a.k.a. free products).
  \item Mention that the subsequent definition can be reduced to a single
    formula (i.e., without cases).
  \item Any measure is a WMC measure if all atoms are in $L$.
  \end{itemize}
}

\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion. Finally, we say that a measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-definable}) if
  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}

\begin{theorem}
  $\WMC$ is a measure, and $\nWMC$ is a probability measure.
\end{theorem}
\begin{proof}
  First, note that $\WMC$ is non-negative and $\WMC(0) = 0$ by definition. Next,
  let $x, y \in \mathbf{B}$ be such that $x \land y = 0$. We want to show that
  \begin{equation} \label{eq:additivity_proof}
    \WMC(x \lor y) = \WMC(x) + \WMC(y).
  \end{equation}
  If, say, $x = 0$, then \cref{eq:additivity_proof} becomes
  \[
    \WMC(y) = \WMC(0) + \WMC(y) = \WMC(y)
  \]
  (and likewise for $y = 0$). Thus we can assume that $x \ne 0 \ne y$ and use
  \cref{thm:representation} to write
  \[
    x = \bigvee_{i \in I} x_i \quad \text{and} \quad y = \bigvee_{j \in J} y_j
  \]
  for some sequences of atoms $(x_i)_{i \in I}$ and $(y_j)_{j \in J}$. If
  $x_{i'} = y_{j'}$ for some $i' \in I$ and $j' \in J$, then
  \[
    x \land y = \bigvee_{i \in I} \bigvee_{j \in J} x_i \land y_j = x_{i'} \land
    y_{j'} \ne 0,
  \]
  contradicting the assumption. This is enough to show that
  \begin{align*}
    \WMC(x \lor y) &= \WMC\left( \left( \bigvee_{i \in I} x_i \right) \lor \left(\bigvee_{j \in J} y_j \right) \right) = \sum_{i \in I} \WMC(x_i) + \sum_{j \in J} \WMC(y_j) \\
                   &= \WMC(x) + \WMC(y),
  \end{align*}
  finishing the proof that $\WMC$ is a measure. This immediately shows that
  $\nWMC$ is a probability measure since, by definition, $\nWMC(1) = 1$.
\end{proof}

@@

\section{Limitations of Literal-Based Weighted Model Counting}

\begin{itemize}
\item[F] Give a concrete example of something impossible to represent using WMC.
\item[F] Can you say something here about factorized vs non-factorized
  weight function definitions? That is, factorized is when $w$ maps literals to
  $R_{\ge 0}$, non-factorized is when $w$ maps models to $R_{\ge 0}$ and
  \begin{itemize}
  \item come up with nice example when non-factorized weights are intuitive;
  \item clarify that the factorized definition have is w.r.t. models, in case some
    one gets confused. [It doesn't have to be, if the BA is not free---P.]
  \end{itemize}
\item What Measures are WMC-Definable: proofs need to be updated and
  propositions could be phrased in a better way, but the gist should be the
  same.
\end{itemize}

\subsection{WMC Requires Independent Literals}

@@

\subsection{Extending the Algebra}

\begin{itemize}
\item[F2] If you prove (b) above, you could motivate why weights on literals
  is attractive and whether there is a way to augment the expressivity of WMC
  while still maintaining literal level weights. Hence this action.
\item[F2] You need to explain the significance of this result.
\end{itemize}

Given this requirement for independence, a well-known way to represent
probability distributions that do not consist entirely of independent variables
is by adding more literals \cite{DBLP:journals/ai/ChaviraD08}, i.e., extending
the set $L$ covered by the WMC weight function $w\colon L \to \mathbb{R}_{\ge
  0}$. Let us translate this idea to the language of BAs.

\begin{theorem} \label{thm:extension}
  Let $\mathbf{B}$ be a free BA over a finite set $S$, and let $m\colon
  \mathbf{B} \to \mathbb{R}_{\ge 0}$ be an arbitrary measure. Let $L = \{ s \mid
  s \in S \} \cup \{ \neg s \mid s \in S \}$. By \cref{lemma:finite_and_free},
  we know that $\mathbf{B}$ has $n = 2^{|S|}$ atoms. Let $\{a_i\}_{i=1}^n$
  denote those atoms in some arbitrary order. Let $L' = L \cup \{ \phi_i
  \}_{i=1}^n \cup \{\neg \phi_i \}_{i=1}^n$ be the set $L$ extended with $2n$
  new elements. Let $\mathbf{B'}$ be the unique Boolean algebra with $\{ \phi_i
  \land a_i \}_{i=1}^n \cup \{ \neg \phi_i \land a_i \}_{i=1}^n$ as its set of
  atoms. Let $\iota\colon \mathbf{B} \hookrightarrow \mathbf{B'}$ be the
  inclusion homomorphism. Let $w\colon L' \to \mathbb{R}_{\ge 0}$ be defined by
  \[
    w(l) = \begin{cases}
      m(a_i)/2 & \text{if } l = \phi_i \text{ or } l = \neg\phi_i \text{ for
        some } i \in [n] \\
      1 & \text{otherwise}
    \end{cases}
  \]
  for all $l \in L'$, and note that this defines a WMC measure $m' =
  \WMC_{w}\colon \mathbf{B'} \to \mathbb{R}_{\ge 0}$. Then
  \[
    m(a) = (m' \circ \iota)(a)
  \]
  for all $a \in \mathbf{B}$.
\end{theorem}

In other words, any measure can be computed using WMC by extending the BA with
more literals. More precisely, we are given the left-hand column in
\[
  \begin{tikzcd}
    \textcolor{red}{\mathbb{R}_{\ge 0}} & & \\
    \textcolor{red}{\mathbf{B}} \arrow[red]{u}{m} \ar[r,hookrightarrow,"\iota"]
    & \mathbf{B'} \arrow{lu}[swap]{m'} & \\
    \textcolor{red}{L} \ar[u,red,hookrightarrow] \ar[r,hookrightarrow] & L'
    \ar[u,hookrightarrow] \arrow{r}{w} & \mathbb{R}_{\ge 0}
  \end{tikzcd}
\]
and construct the remaining part in such a way that the triangle commutes.

\todo[inline,caption={}]{
  \begin{itemize}
  \item Make $J$ depend on $i$.
  \item Find a reference for this first claim in the following proof.
  \end{itemize}
}
\begin{proof}
  Since $\mathbf{B}$ is free over $S$, each atom $a_i \in \mathbf{B}$ is an
  infimum of elements in $L$, i.e.,
  \[
    a_i = \bigwedge_{j \in J} a_{i,j}
  \]
  for some $\{ a_{i,j} \}_{j \in J} \subset L$. Moreover, each atom $b \in
  \mathbf{B'}$ can be represented as either $b = \phi_i \land a_i$ or $b =
  \neg\phi_i \land a_i$ for some atom $a_i \in \mathbf{B}$, also making it an
  infimum over a subset of $L'$. Then, for any $b \in \mathbf{B}$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) \prod_{j \in
    J} w(a_{i,j}),
  \]
  recognising that, for any $\iota(b)$, any atom $a_i \in \mathbf{B}$ satisfies
  $\phi_i \land a_i \le \iota(b)$ if and only if it satisfies $\neg\phi_i \land
  a_i \le \iota(b)$. Then, according to the definition of $w$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) =
    \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\ \phi_i \land a_i \le
        \iota(b)}} m(a_i) = m(b),
  \]
  provided that
  \[
    \phi_i \land a_i \le \iota(b) \quad \text{if and only if} \quad a_i \le b,
  \]
  but this is equivalent to
  \[
    \phi_i \land a_i = \phi_i \land a_i \land b \quad \text{if and only if}
    \quad a_i = a_i \land b
  \]
  which is true because $\phi_i \not\in L$.
\end{proof}

Now we can show that the construction in \cref{thm:extension} is smallest
possible.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}
\usepackage{multirow}

\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

\begin{itemize}
\item The Main Narrative
  \begin{enumerate}
  \item When weights are defined on literals, the measure on the free BA is
    fully independent.
  \item This means that the BA itself must be larger (i.e., have additional
    `meaningless' literals) to turn any probability distribution into an
    independent one.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any probability distribution into a Boolean algebra that's not
    necessarily independent and thus can be smaller.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs, and thus a simpler problem.
  \item (Maybe) we experimentally demonstrate a phase transition based on the
    number of variables per ADD.
  \end{enumerate}
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item Intuitively, a measure is just like a probability, except it's in
  $\mathbb{R}_{\ge 0}$ instead of $[0, 1]$.
\item ADDMC should probably be in textsc mode.
\end{itemize}

@@

\begin{itemize}
\item Make up my mind about $a, b$ vs. $x, y$ and stick to it (maybe $x, y$?).
\item Terminology: `with generating set $S$' $\to$ `over $S$'.
\item Notation: if $L$ denotes literals, then it doesn't denote a generating
  set. Literals should be $U$.
\end{itemize}

\begin{table}
  \caption{(in the same order)}
  \centering
  \begin{tabular}{lcc}
    \toprule
    & Set-theoretic notation & Boolean-algebraic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\emptyset, \{a\}, \{b\}, \{a, b\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ & $\top$ \\
    & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ & $\neg a \lor \neg b, a \to b$ \\
    & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ & $b \to a, a \lor b$ \\
    & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ & $\neg b, \neg a, a \leftrightarrow b$ \\
    & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ & $(a \land \neg b) \lor (b \land \neg a), a, b$ \\
    & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\emptyset$ & $\bot$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

@@

\subsection{The Space of Functions on Boolean Algebras}

\begin{definition}[Operations on functions]
  Let $A\colon 2^X \to \mathbb{R}_{\ge 0}$ and $B\colon 2^Y \to \mathbb{R}_{\ge
    0}$ be Boolean functions, $\alpha \in \mathbb{R}_{\ge 0}$, and $x \in X$. We
  define the following operations:
  \begin{description}
  \item[Addition:] $A+B$ is a function $A+B\colon 2^{X \cup Y} \to
    \mathbb{R}_{\ge 0}$ such that
    \[
      (A+B)(\tau) = A(\tau \cap X) + B(\tau \cap Y)
    \]
    for all $\tau \in 2^{X \cup Y}$.
  \item[Inverse:] $\overline{A}$ is a function $\overline{A}\colon 2^X \to
    \mathbb{R}_{\ge 0}$ such that
    \[
      \overline{A}(\tau) = 1 - A(\tau)
    \]
    for all $\tau \in 2^X$.
  \item[Multiplication:] $A \cdot B$ is a function $A \cdot B\colon 2^{X \cup Y}
    \to \mathbb{R}_{\ge 0}$ such that
    \[
      (A \cdot B)(\tau) = A(\tau \cap X) \cdot B(\tau \cap Y)
    \]
    for all $\tau \in 2^{X \cup Y}$.
  \item[Scalar multiplication:] $\alpha A$ is a function $\alpha A\colon 2^X \to
    \mathbb{R}_{\ge 0}$ such that
    \[
      (\alpha A)(\tau) = \alpha \cdot A(\tau)
    \]
    for all $\tau \in 2^X$.
  \item[Projection:] $\exists_xA$ is a function $\exists_xA\colon 2^{X \setminus
      \{ x \}} \to \mathbb{R}_{\ge 0}$ such that
    \[
      (\exists_xA)(\tau) = A(\tau) + A(\tau \cup \{ x \})
    \]
    for all $\tau \in 2^{X \setminus \{x \}}$.
  \end{description}
\end{definition}

\begin{observation}
  Let $\mathcal{V} = \{ A\colon 2^X \to \mathbb{R}_{\ge 0} \mid X \subseteq U
  \}$. Then $\mathcal{V}$ is a semi-vector space with three additional
  operations: inverse, (non-scalar) multiplication, and projection.
  Specifically, note that both addition and multiplication are both associative
  and commutative.
\end{observation}

@@

\begin{itemize}
\item[F2] Explain with examples: models $=$ elements [atoms] of algebra.
\item Be careful about mentioning ideals and quotients.
\item Note that we're going to use logical notation for BAs. Give an example of
  the two notations.
\end{itemize}

\begin{example} \label{example:construction}
  % TODO: keep this, but might need a lot of editing.
  Let $\mathcal{L}$ be a propositional logic with $p$ and $q$ as its only atoms.
  Then $L = \{ p, q, \neg p, \neg q \}$ is its set of literals. Let $w : L \to
  \mathbb{R}_{\ge 0}$ be the \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \\
    w(\neg p) = 0.7, \\
    w(q) = 0.2, \\
    w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of $\Delta$ is
  then
  \[
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \]

  The corresponding BA $B(\Delta)$ can then be constructed as a
  Lindenbaum-Tarski algebra. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$ and then take a quotient with respect to
  either the filter generated by $p$ or the ideal\footnote{More details on these
    concepts can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate}
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
\todo[inline,caption={}]{
  \begin{itemize}
  \item Why is Step 1 always possible?
  \item Clarify what $B(L)$ means and whether $B(\Delta)$ is even necessary.
  \item Find a reference for the set/subset thing.
  \end{itemize}
}

\begin{definition} \label{def:wmc}
  Let $U$ be an arbitrary set. Then
  \begin{itemize}
  \item a \emph{measure} is a function $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$,
  \item a \emph{weight function} is a function $W\colon 2^U \to \mathbb{R}_{\ge
      0}$.
  \item a weight function is \emph{factored} if $W = \prod_{x \in U} W_x$ for
    some functions $W_x\colon 2^{\{x\}} \to \mathbb{R}_{\ge 0}$.
  \item every weight function induces a measure:
    \[
      M_W(x) = \begin{cases}
        0 & \text{if } x = \emptyset \\
        W(u) & \text{if } x = \{ u \} \\
        \sum_{\{u\} \le x} W(u) & \text{otherwise.}
      \end{cases}
    \]
    The process of calculating the value of $M_W(x)$ for some $x \in 2^{2^U}$
    with a given definition of $W$ is known as \emph{weighted model counting}.
  \end{itemize}
\end{definition}

@@

\section{Limitations of Factored Weighted Model Counting}

\begin{itemize}
\item[F] Give a concrete example of something impossible to represent using WMC.
\item[F] Can you say something here about factorized vs non-factorized
  weight function definitions? That is, factorized is when $w$ maps literals to
  $R_{\ge 0}$, non-factorized is when $w$ maps models to $R_{\ge 0}$ and
  \begin{itemize}
  \item come up with nice example when non-factorized weights are intuitive;
  \item clarify that the factorized definition have is w.r.t. models, in case some
    one gets confused. [It doesn't have to be, if the BA is not free---P.]
  \end{itemize}
\end{itemize}

@@

Given this requirement for independence, a well-known way to represent
probability distributions that do not consist entirely of independent variables
is by adding more literals \cite{DBLP:journals/ai/ChaviraD08}, i.e., extending
the set $L$ covered by the WMC weight function $w\colon L \to \mathbb{R}_{\ge
  0}$. More precisely, we are given the left-hand column in
\[
  \begin{tikzcd}
    \textcolor{red}{\mathbb{R}_{\ge 0}} & & \\
    \textcolor{red}{\mathbf{B}} \arrow[red]{u}{m} \ar[r,hookrightarrow,"\iota"]
    & \mathbf{B'} \arrow{lu}[swap]{m'} & \\
    \textcolor{red}{L} \ar[u,red,hookrightarrow] \ar[r,hookrightarrow] & L'
    \ar[u,hookrightarrow] \arrow{r}{w} & \mathbb{R}_{\ge 0}
  \end{tikzcd}
\]
and construct the remaining part in such a way that the triangle commutes.

\todo[inline]{Must quotient the BA w.r.t. some rules.}
